SunPy - Python for Solar Physics

### 2014-01-27 3-2-HEK.tex

Simple HEK queries consist of start time, an end time, and an event
type (Example \ref{code:hek:simple}). Times can be specified in any
SunPy-acceptable format. Event types are specified as upper case, two
letter strings, and are identical to the two letter abbreviations
found at the HEK website, \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

\begin{listing}[H]
\begin{minted}{python}
from sunpy.net import hek
client = hek.HEKClient()
tstart = '2011/08/09 07:23:56'
tend = '2011/08/09 12:40:29'
event\_type = 'FL'
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type))
\end{minted}
\caption{Simple HEK query, returning 19 entries.}
\label{code:hek:simple}
\end{listing}

The module \texttt{hek.attrs} contains attributes of the HEK that can be used to
construct HEK queries.  For example, a flare is an attribute of the
HEK, and so instead of specifying \texttt{hek.attrs.EventType('FL')}
in Example \ref{code:hek:simple}, this can also be expressed as
\texttt{hek.attrs.FL}.

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in
that many of them are wrappers that conveniently expose
comparisions by overloading Python operators.  This allows filtering
of the HEK entries by the properties of the event.  As was mentioned
above, the HEK stores feature/event metadata obtained in different
ways, known generally as {\it feature recognition methods}, or FRMs.
Example \ref{code:hek:frm} filters the results of the previous result
to return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type), hek.attrs.FRM.Name == 'SSW Latest Events')
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators AND
and OR. This makes complex queries easy to create: example
\ref{code:hek:or} returns flares west of 50 arcseconds or those that
have a peak flux above 1000.0 (the units of the flux are FRM dependent
and are described at the HEK website).
\begin{listing}[H]
\begin{minted}{python}
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event\_type), (hek.attrs.Event.Coord1 > 50) or 
(hek.attrs.FL.PeakFlux > 1000.0) )
\end{minted}
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}), but the optional
attributes are FRM dependent.  If a FRM does not have one of the
optional attribute, a 'None' is returned by the \texttt{hek} module.

The ability to use comparison and logical operators on HEK attributes
allows the construction of queries of almost arbitrary complexity,
bringing powerful feature/event search capabilities to SunPy.

>>>>

Simple HEK queries consist of start time, an end time, and an event type 
(Listing \ref{code:hek:simple}). Times can be specified in any SunPy-acceptable 
format. Event types are specified as upper case, two letter strings, and are 
identical to the two letter abbreviations found at the HEK website, 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
from sunpy.net import hek
client = hek.HEKClient()
tstart = '2011/08/09 07:23:56'
tend = '2011/08/09 12:40:29'
event_type = 'FL'
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event_type))
\end{minted}
\caption{Simple HEK query, returning 19 entries.}
\label{code:hek:simple}
\end{listing}

The module \texttt{hek.attrs} contains attributes of the HEK that can be used to
construct HEK queries.  For example, a flare is an attribute of the HEK, and so 
instead of specifying \texttt{hek.attrs.EventType('FL')} in Listing 
\ref{code:hek:simple}, this can also be expressed as \texttt{hek.attrs.FL}.

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods}, or 
FRMs. Example \ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
                      hek.attrs.FRM.Name == 'SSW Latest Events')
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators e.g. AND
and OR. This makes complex queries easy to create: Listing \ref{code:hek:or} 
returns flares west of 50 arcseconds or those that have a peak flux above 
1000.0 (the units of the flux are FRM dependent and are described at the HEK 
website).
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
      (hek.attrs.Event.Coord1 > 50) or (hek.attrs.FL.PeakFlux > 1000.0))
\end{minted}
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}), but the optional attributes 
are FRM dependent.  If a FRM does not have one of the optional attribute, a 
'None' is returned by the \texttt{hek} module.

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity, bringing powerful 
feature/event search capabilities to SunPy. It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.
